### 2022-11-29 paper/annotated_bibliography.tex
move all paper files to directory

\documentclass[11pt,reqno]{article}
%\usepackage{cases}

\RequirePackage{amsthm,amsmath,amsfonts,amssymb}
%\RequirePackage[numbers]{natbib}
\RequirePackage[authoryear]{natbib}%% uncomment this for author-year citations
\RequirePackage[colorlinks,citecolor=blue,urlcolor=blue]{hyperref}%% uncomment this for coloring bibliography citations and linked URLs
\RequirePackage{graphicx}%% uncomment this for including figures

\usepackage{amsmath, amssymb, amscd, amsthm, amsfonts}

\begin{document}

\title{Annotated Bibliography}
\author{Rob Zielinski}

\maketitle

\nocite{*}
\bibliographystyle{plain-annote}
\bibliography{references}

\end{document}