### 2024-07-10 code/05_lpme_adni_results.qmd

---
title: "ADNI Application Results"
author: "Robert Zielinski"
date: "June 14, 2024"
output: html_document

execute:
  eval: false
---

This notebook includes the code used to generate plots and other visualizations from the results of applying the LPME and PME algorithms to ADNI data, as well as the code needed to summarize the results as shown in the paper.

```{r}
library(bazar)
library(brolgar)
library(here)
library(numbers)
library(plot3D)
library(plotly)
library(RColorBrewer)
library(scatterplot3d)
library(tidyverse)

source(here("code/functions/create_cross_section_matrix.R"))
```

```{r}
# set color palette to maintain consistency between plots
colors <- brewer.pal(3, "Set1")
```

```{r}
result_dirs <- list.dirs(here("results"), recursive = FALSE)
```

```{r}
lhipp_full <- read_csv(here("data/adni_lhipp_surface_processed.csv"))
rhipp_full <- read_csv(here("data/adni_rhipp_surface_processed.csv"))
lthal_full <- read_csv(here("data/adni_lthal_surface_processed.csv"))
rthal_full <- read_csv(here("data/adni_rthal_surface_processed.csv"))

patnos <- unique(lhipp_full$patno)
```

```{r}
for (dir_idx in 1:length(result_dirs)) {
  dir_path <- result_dirs[dir_idx]

  # load files with structure surface data
  lhipp_mat <- read_csv(paste0(dir_path, "/adni_lhipp_mat.csv"))[, -1]
  rhipp_mat <- read_csv(paste0(dir_path, "/adni_rhipp_mat.csv"))[, -1]
  lthal_mat <- read_csv(paste0(dir_path, "/adni_lthal_mat.csv"))[, -1]
  rthal_mat <- read_csv(paste0(dir_path, "/adni_rthal_mat.csv"))[, -1]

  lhipp <- lhipp_full %>% 
    filter(patno == patnos[dir_idx])
  rhipp <- rhipp_full %>% 
    filter(patno == patnos[dir_idx])
  lthal <- lthal_full %>%
    filter(patno == patnos[dir_idx])
  rthal <- rthal_full %>%
    filter(patno == patnos[dir_idx])

  lhipp_mat <- lhipp %>% 
    select(time_from_bl, x, y, z, theta, phi, r) %>% 
    as.matrix()
  rhipp_mat <- rhipp %>% 
    select(time_from_bl, x, y, z, theta, phi, r) %>% 
    as.matrix()
  lthal_mat <- lthal %>% 
    select(time_from_bl, x, y, z, theta, phi, r) %>% 
    as.matrix()
  rthal_mat <- rthal %>% 
    select(time_from_bl, x, y, z, theta, phi, r) %>% 
    as.matrix()

  # load lpme and pme estimates of all structures
  lhipp_lpme_isomap_vals <- read_csv(paste0(dir_path, "/adni_lhipp_lpme_isomap_vals.csv"))[, -1] %>% 
    as.matrix()
  lhipp_pme_vals <- read_csv(paste0(dir_path, "/adni_lhipp_pme_vals.csv"))[, -1] %>% 
    as.matrix()
  rhipp_lpme_isomap_vals <- read_csv(paste0(dir_path, "/adni_rhipp_lpme_isomap_vals.csv"))[, -1] %>% 
    as.matrix()
  rhipp_pme_vals <- read_csv(paste0(dir_path, "/adni_rhipp_pme_vals.csv"))[, -1] %>% 
    as.matrix()
  lthal_lpme_isomap_vals <- read_csv(paste0(dir_path, "/adni_lthal_lpme_isomap_vals.csv"))[, -1] %>% 
    as.matrix()
  lthal_pme_vals <- read_csv(paste0(dir_path, "/adni_lthal_pme_vals.csv"))[, -1] %>% 
    as.matrix()
  rthal_lpme_isomap_vals <- read_csv(paste0(dir_path, "/adni_rthal_lpme_isomap_vals.csv"))[, -1] %>% 
    as.matrix()
  rthal_pme_vals <- read_csv(paste0(dir_path, "/adni_rthal_pme_vals.csv"))[, -1] %>% 
    as.matrix()

  time_vals <- unlist(lhipp_mat[, 1]) %>% 
    unique()

  nplots <- floor(length(time_vals) / 2)

  lhipp_range <- colMinsMaxs(lhipp_mat[, 2:4]) %>% 
    abs() %>% 
    max()
  rhipp_range <- colMinsMaxs(rhipp_mat[, 2:4]) %>% 
    abs() %>% 
    max()
  lthal_range <- colMinsMaxs(lthal_mat[, 2:4]) %>% 
    abs() %>% 
    max()
  rthal_range <- colMinsMaxs(rthal_mat[, 2:4]) %>% 
    abs() %>% 
    max()

  lhipp_positions <- data.frame(
    c(0, round(-1.35 * lhipp_range, 2), round(1.35 * lhipp_range, 2)),
    c(round(1.45 * lhipp_range, 2), 0, round(1.35 * lhipp_range, 2)),
    c(round(-1.2 * lhipp_range, 2), round(-1.3 * lhipp_range, 2), 0)
  ) %>% 
    as.matrix(byrow = TRUE)

  rhipp_positions <- data.frame(
    c(0, round(-1.35 * rhipp_range, 2), round(1.35 * rhipp_range, 2)),
    c(round(1.45 * rhipp_range, 2), 0, round(1.35 * rhipp_range, 2)),
    c(round(-1.2 * rhipp_range, 2), round(-1.3 * rhipp_range, 2), 0)
  ) %>% 
    as.matrix(byrow = TRUE)

  lthal_positions <- data.frame(
    c(0, round(-1.35 * lthal_range, 2), round(1.35 * lthal_range, 2)),
    c(round(1.45 * lthal_range, 2), 0, round(1.35 * lthal_range, 2)),
    c(round(-1.2 * lthal_range, 2), round(-1.3 * lthal_range, 2), 0)
  ) %>% 
    as.matrix(byrow = TRUE)

  rthal_positions <- data.frame(
    c(0, round(-1.35 * rthal_range, 2), round(1.35 * rthal_range, 2)),
    c(round(1.45 * rthal_range, 2), 0, round(1.35 * rthal_range, 2)),
    c(round(-1.2 * rthal_range, 2), round(-1.3 * rthal_range, 2), 0)
  ) %>% 
    as.matrix(byrow = TRUE)

  plot_labels <- c("x", "y", "z")

  p <- plot_ly(
    x = lhipp_mat[, 2],
    y = lhipp_mat[, 3],
    z = lhipp_mat[, 4],
    frame = lhipp_mat[, 1],
    type = "scatter3d",
    mode = "markers",
    marker = list(size = 3),
    name = "Data"
  ) %>%
    add_markers(
      x = lhipp_lpme_isomap_vals[, 2],
      y = lhipp_lpme_isomap_vals[, 3],
      z = lhipp_lpme_isomap_vals[, 4],
      frame = lhipp_lpme_isomap_vals[, 1],
      name = "LPME"
    ) %>%
    add_markers(
      x = lhipp_pme_vals[, 2],
      y = lhipp_pme_vals[, 3],
      z = lhipp_pme_vals[, 4],
      frame = lhipp_pme_vals[, 1],
      name = "PME"
    )

  lhipp_lpme_plots <- list()
  lhipp_pme_plots <- list()
  lhipp_data_plots <- list()

  lims <- c(-lhipp_range, lhipp_range)
  png(
    paste0(dir_path, "/adni_lhipp_data_plot.png"),
    res = 1000,
    height = 3500,
    width = 5000 * nplots
  )
  par(
    oma = c(0, 0, 1, 0), 
    mfrow = c(1, floor(length(time_vals) / 2)), 
    mar = c(0, 0, 1, 0), 
    mgp = c(5, 1, 0)
  )

  for (t in 1:length(time_vals)) {
    if (mod(t, 2) > 1e-10) {
      next
    } else {
      temp_data <- lhipp_mat[almost.equal(lhipp_mat[, 1], time_vals[t]), ]

      scatter3D(
        x = temp_data[, 2],
        y = temp_data[, 3],
        z = temp_data[, 4],
        pch = 20,
        col = alpha(colors[1], 0.3),
        theta = 220,
        ticktype = "detailed",
        nticks = 1,
        main = paste0("Time = ", round((time_vals)[t], 2)),
        xlim = lims,
        ylim = lims,
        zlim = lims,
        xlab = "",
        ylab = "",
        zlab = "",
        cex.axis = 0.67,
        cex = 0.75
      )
      text3D(
        x = lhipp_positions[, 1],
        y = lhipp_positions[, 2],
        z = lhipp_positions[, 3],
        labels = plot_labels,
        add = TRUE
      )
    }
  }
  dev.off()

  png(
    paste0(dir_path, "/adni_lhipp_lpme_isomap_plot.png"),
    res = 1000,
    height = 3500,
    width = 5000 * nplots
  )
  par(
    oma = c(0, 0, 1, 0), 
    mfrow = c(1, floor(length(time_vals) / 2)), 
    mar = c(0, 0, 1, 0), 
    mgp = c(5, 1, 0)
  )

  for (t in 1:length(time_vals)) {
    if (mod(t, 2) > 1e-10) {
      next
    } else {
      temp_lpme <- lhipp_lpme_isomap_vals[almost.equal(lhipp_lpme_isomap_vals[, 1], time_vals[t]), ]
      scatter3D(
        x = temp_lpme[, 2],
        y = temp_lpme[, 3],
        z = temp_lpme[, 4],
        pch = 15,
        col = alpha(colors[2], 0.3),
        theta = 220,
        ticktype = "detailed",
        nticks = 1,
        main = paste0("Time = ", round((time_vals)[t], 2)),
        xlim = lims,
        ylim = lims,
        zlim = lims,
        xlab = "",
        ylab = "",
        zlab = "",
        cex.axis = 0.67,
        cex = 0.75
      )
      text3D(
        x = lhipp_positions[, 1],
        y = lhipp_positions[, 2],
        z = lhipp_positions[, 3],
        labels = plot_labels,
        add = TRUE
      )
    }
  }
  dev.off()

  png(
    paste0(dir_path, "/adni_lhipp_pme_plot.png"),
    res = 1000,
    height = 3500,
    width = 5000 * nplots
  )
  par(
    oma = c(0, 0, 1, 0), 
    mfrow = c(1, floor(length(time_vals) / 2)), 
    mar = c(0, 0, 1, 0), 
    mgp = c(5, 1, 0)
  )

  for (t in 1:length(time_vals)) {
    if (mod(t, 2) > 1e-10) {
      next
    } else {
      temp_pme <- lhipp_pme_vals[almost.equal(lhipp_pme_vals[, 1], time_vals[t]), ]
      scatter3D(
        x = temp_pme[, 2],
        y = temp_pme[, 3],
        z = temp_pme[, 4],
        pch = 17,
        col = alpha(colors[3], 0.3),
        theta = 220,
        ticktype = "detailed",
        nticks = 1,
        main = paste0("Time = ", round((time_vals)[t], 2)),
        xlim = lims,
        ylim = lims,
        zlim = lims,
        xlab = "",
        ylab = "",
        zlab = "",
        cex.axis = 0.67,
        cex = 0.75
      )
      text3D(
        x = lhipp_positions[, 1],
        y = lhipp_positions[, 2],
        z = lhipp_positions[, 3],
        labels = plot_labels,
        add = TRUE
      )
    }
  }
  dev.off()

  lhipp_cross_section <- create_cross_section_matrix(
    list(lhipp_mat, lhipp_lpme_isomap_vals, lhipp_pme_vals),
    4,
    c(0.005, 0.005, 0.005),
    nrow = 2,
    lhipp = lhipp
  )
  ggsave(
    paste0(dir_path, "/adni_lhipp_cross_section.png"),
    plot = lhipp_cross_section,
    height = 5000,
    width = 7000,
    units = "px",
    dpi = 1250
  )

  lims <- c(-rhipp_range, rhipp_range)
  png(
    paste0(dir_path, "/adni_rhipp_data_plot.png"),
    res = 1000,
    height = 3500,
    width = 5000 * nplots
  )
  par(
    oma = c(0, 0, 1, 0), 
    mfrow = c(1, floor(length(time_vals) / 2)), 
    mar = c(0, 0, 1, 0), 
    mgp = c(5, 1, 0)
  )

  for (t in 1:length(time_vals)) {
    if (mod(t, 2) > 1e-10) {
      next
    } else {
      temp_data <- rhipp_mat[almost.equal(rhipp_mat[, 1], time_vals[t]), ]

      scatter3D(
        x = temp_data[, 2],
        y = temp_data[, 3],
        z = temp_data[, 4],
        pch = 20,
        col = alpha(colors[1], 0.3),
        theta = 220,
        ticktype = "detailed",
        nticks = 1,
        main = paste0("Time = ", round((time_vals)[t], 2)),
        xlim = lims,
        ylim = lims,
        zlim = lims,
        xlab = "",
        ylab = "",
        zlab = "",
        cex.axis = 0.67,
        cex = 0.75
      )
      text3D(
        x = rhipp_positions[, 1],
        y = rhipp_positions[, 2],
        z = rhipp_positions[, 3],
        labels = plot_labels,
        add = TRUE
      )
    }
  }
  dev.off()

  png(
    paste0(dir_path, "/adni_rhipp_lpme_isomap_plot.png"),
    res = 1000,
    height = 3500,
    width = 5000 * nplots
  )
  par(
    oma = c(0, 0, 1, 0), 
    mfrow = c(1, floor(length(time_vals) / 2)), 
    mar = c(0, 0, 1, 0), 
    mgp = c(5, 1, 0)
  )

  for (t in 1:length(time_vals)) {
    if (mod(t, 2) > 1e-10) {
      next
    } else {
      temp_lpme <- rhipp_lpme_isomap_vals[almost.equal(rhipp_lpme_isomap_vals[, 1], time_vals[t]), ]
      scatter3D(
        x = temp_lpme[, 2],
        y = temp_lpme[, 3],
        z = temp_lpme[, 4],
        pch = 15,
        col = alpha(colors[2], 0.3),
        theta = 220,
        ticktype = "detailed",
        nticks = 1,
        main = paste0("Time = ", round((time_vals)[t], 2)),
        xlim = lims,
        ylim = lims,
        zlim = lims,
        xlab = "",
        ylab = "",
        zlab = "",
        cex.axis = 0.67,
        cex = 0.75
      )
      text3D(
        x = rhipp_positions[, 1],
        y = rhipp_positions[, 2],
        z = rhipp_positions[, 3],
        labels = plot_labels,
        add = TRUE
      )
    }
  }
  dev.off()

  png(
    paste0(dir_path, "/adni_rhipp_pme_plot.png"),
    res = 1000,
    height = 3500,
    width = 5000 * nplots
  )
  par(
    oma = c(0, 0, 1, 0), 
    mfrow = c(1, floor(length(time_vals) / 2)), 
    mar = c(0, 0, 1, 0), 
    mgp = c(5, 1, 0)
  )

  for (t in 1:length(time_vals)) {
    if (mod(t, 2) > 1e-10) {
      next
    } else {
      temp_pme <- rhipp_pme_vals[almost.equal(rhipp_pme_vals[, 1], time_vals[t]), ]
      scatter3D(
        x = temp_pme[, 2],
        y = temp_pme[, 3],
        z = temp_pme[, 4],
        pch = 17,
        col = alpha(colors[3], 0.3),
        theta = 220,
        ticktype = "detailed",
        nticks = 1,
        main = paste0("Time = ", round((time_vals)[t], 2)),
        xlim = lims,
        ylim = lims,
        zlim = lims,
        xlab = "",
        ylab = "",
        zlab = "",
        cex.axis = 0.67,
        cex = 0.75
      )
      text3D(
        x = rhipp_positions[, 1],
        y = rhipp_positions[, 2],
        z = rhipp_positions[, 3],
        labels = plot_labels,
        add = TRUE
      )
    }
  }
  dev.off()

  rhipp_cross_section <- create_cross_section_matrix(
    list(rhipp_mat, rhipp_lpme_isomap_vals, rhipp_pme_vals),
    4,
    c(0.005, 0.005, 0.005),
    nrow = 2,
    lhipp = rhipp
  )
  ggsave(
    paste0(dir_path, "/adni_rhipp_cross_section.png"),
    plot = rhipp_cross_section,
    height = 5000,
    width = 7000,
    units = "px",
    dpi = 1250
  )

  lims <- c(-lthal_range, lthal_range)
  png(
    paste0(dir_path, "/adni_lthal_data_plot.png"),
    res = 1000,
    height = 3500,
    width = 5000 * nplots
  )
  par(
    oma = c(0, 0, 1, 0), 
    mfrow = c(1, floor(length(time_vals) / 2)), 
    mar = c(0, 0, 1, 0), 
    mgp = c(5, 1, 0)
  )

  for (t in 1:length(time_vals)) {
    if (mod(t, 2) > 1e-10) {
      next
    } else {
      temp_data <- lthal_mat[almost.equal(lthal_mat[, 1], time_vals[t]), ]

      scatter3D(
        x = temp_data[, 2],
        y = temp_data[, 3],
        z = temp_data[, 4],
        pch = 20,
        col = alpha(colors[1], 0.3),
        theta = 220,
        ticktype = "detailed",
        nticks = 1,
        main = paste0("Time = ", round((time_vals)[t], 2)),
        xlim = lims,
        ylim = lims,
        zlim = lims,
        xlab = "",
        ylab = "",
        zlab = "",
        cex.axis = 0.67,
        cex = 0.75
      )
      text3D(
        x = lthal_positions[, 1],
        y = lthal_positions[, 2],
        z = lthal_positions[, 3],
        labels = plot_labels,
        add = TRUE
      )
    }
  }
  dev.off()

  png(
    paste0(dir_path, "/adni_lthal_lpme_isomap_plot.png"),
    res = 1000,
    height = 3500,
    width = 5000 * nplots
  )
  par(
    oma = c(0, 0, 1, 0), 
    mfrow = c(1, floor(length(time_vals) / 2)), 
    mar = c(0, 0, 1, 0), 
    mgp = c(5, 1, 0)
  )

  for (t in 1:length(time_vals)) {
    if (mod(t, 2) > 1e-10) {
      next
    } else {
      temp_lpme <- lthal_lpme_isomap_vals[almost.equal(lthal_lpme_isomap_vals[, 1], time_vals[t]), ]
      scatter3D(
        x = temp_lpme[, 2],
        y = temp_lpme[, 3],
        z = temp_lpme[, 4],
        pch = 15,
        col = alpha(colors[2], 0.3),
        theta = 220,
        ticktype = "detailed",
        nticks = 1,
        main = paste0("Time = ", round((time_vals)[t], 2)),
        xlim = lims,
        ylim = lims,
        zlim = lims,
        xlab = "",
        ylab = "",
        zlab = "",
        cex.axis = 0.67,
        cex = 0.75
      )
      text3D(
        x = lthal_positions[, 1],
        y = lthal_positions[, 2],
        z = lthal_positions[, 3],
        labels = plot_labels,
        add = TRUE
      )
    }
  }
  dev.off()

  png(
    paste0(dir_path, "/adni_lthal_pme_plot.png"),
    res = 1000,
    height = 3500,
    width = 5000 * nplots
  )
  par(
    oma = c(0, 0, 1, 0), 
    mfrow = c(1, floor(length(time_vals) / 2)), 
    mar = c(0, 0, 1, 0), 
    mgp = c(5, 1, 0)
  )

  for (t in 1:length(time_vals)) {
    if (mod(t, 2) > 1e-10) {
      next
    } else {
      temp_pme <- lthal_pme_vals[almost.equal(lthal_pme_vals[, 1], time_vals[t]), ]
      scatter3D(
        x = temp_pme[, 2],
        y = temp_pme[, 3],
        z = temp_pme[, 4],
        pch = 17,
        col = alpha(colors[3], 0.3),
        theta = 220,
        ticktype = "detailed",
        nticks = 1,
        main = paste0("Time = ", round((time_vals)[t], 2)),
        xlim = lims,
        ylim = lims,
        zlim = lims,
        xlab = "",
        ylab = "",
        zlab = "",
        cex.axis = 0.67,
        cex = 0.75
      )
      text3D(
        x = lthal_positions[, 1],
        y = lthal_positions[, 2],
        z = lthal_positions[, 3],
        labels = plot_labels,
        add = TRUE
      )
    }
  }
  dev.off()

  lthal_cross_section <- create_cross_section_matrix(
    list(lthal_mat, lthal_lpme_isomap_vals, lthal_pme_vals),
    4,
    c(0.005, 0.005, 0.005),
    nrow = 2,
    lhipp = lthal
  )
  ggsave(
    paste0(dir_path, "/adni_lthal_cross_section.png"),
    plot = lthal_cross_section,
    height = 5000,
    width = 7000,
    units = "px",
    dpi = 1250
  )

  lims <- c(-rthal_range, rthal_range)
  png(
    paste0(dir_path, "/adni_rthal_data_plot.png"),
    res = 1000,
    height = 3500,
    width = 5000 * nplots
  )
  par(
    oma = c(0, 0, 1, 0), 
    mfrow = c(1, floor(length(time_vals) / 2)), 
    mar = c(0, 0, 1, 0), 
    mgp = c(5, 1, 0)
  )

  for (t in 1:length(time_vals)) {
    if (mod(t, 2) > 1e-10) {
      next
    } else {
      temp_data <- rthal_mat[almost.equal(rthal_mat[, 1], time_vals[t]), ]

      scatter3D(
        x = temp_data[, 2],
        y = temp_data[, 3],
        z = temp_data[, 4],
        pch = 20,
        col = alpha(colors[1], 0.3),
        theta = 220,
        ticktype = "detailed",
        nticks = 1,
        main = paste0("Time = ", round((time_vals)[t], 2)),
        xlim = lims,
        ylim = lims,
        zlim = lims,
        xlab = "",
        ylab = "",
        zlab = "",
        cex.axis = 0.67,
        cex = 0.75
      )
      text3D(
        x = rthal_positions[, 1],
        y = rthal_positions[, 2],
        z = rthal_positions[, 3],
        labels = plot_labels,
        add = TRUE
      )
    }
  }
  dev.off()

  png(
    paste0(dir_path, "/adni_rthal_lpme_isomap_plot.png"),
    res = 1000,
    height = 3500,
    width = 5000 * nplots
  )
  par(
    oma = c(0, 0, 1, 0), 
    mfrow = c(1, floor(length(time_vals) / 2)), 
    mar = c(0, 0, 1, 0), 
    mgp = c(5, 1, 0)
  )

  for (t in 1:length(time_vals)) {
    if (mod(t, 2) > 1e-10) {
      next
    } else {
      temp_lpme <- rthal_lpme_isomap_vals[almost.equal(rthal_lpme_isomap_vals[, 1], time_vals[t]), ]
      scatter3D(
        x = temp_lpme[, 2],
        y = temp_lpme[, 3],
        z = temp_lpme[, 4],
        pch = 15,
        col = alpha(colors[2], 0.3),
        theta = 220,
        ticktype = "detailed",
        nticks = 1,
        main = paste0("Time = ", round((time_vals)[t], 2)),
        xlim = lims,
        ylim = lims,
        zlim = lims,
        xlab = "",
        ylab = "",
        zlab = "",
        cex.axis = 0.67,
        cex = 0.75
      )
      text3D(
        x = rthal_positions[, 1],
        y = rthal_positions[, 2],
        z = rthal_positions[, 3],
        labels = plot_labels,
        add = TRUE
      )
    }
  }
  dev.off()

  png(
    paste0(dir_path, "/adni_rthal_pme_plot.png"),
    res = 1000,
    height = 3500,
    width = 5000 * nplots
  )
  par(
    oma = c(0, 0, 1, 0), 
    mfrow = c(1, floor(length(time_vals) / 2)), 
    mar = c(0, 0, 1, 0), 
    mgp = c(5, 1, 0)
  )

  for (t in 1:length(time_vals)) {
    if (mod(t, 2) > 1e-10) {
      next
    } else {
      temp_pme <- rthal_pme_vals[almost.equal(rthal_pme_vals[, 1], time_vals[t]), ]
      scatter3D(
        x = temp_pme[, 2],
        y = temp_pme[, 3],
        z = temp_pme[, 4],
        pch = 17,
        col = alpha(colors[3], 0.3),
        theta = 220,
        ticktype = "detailed",
        nticks = 1,
        main = paste0("Time = ", round((time_vals)[t], 2)),
        xlim = lims,
        ylim = lims,
        zlim = lims,
        xlab = "",
        ylab = "",
        zlab = "",
        cex.axis = 0.67,
        cex = 0.75
      )
      text3D(
        x = rthal_positions[, 1],
        y = rthal_positions[, 2],
        z = rthal_positions[, 3],
        labels = plot_labels,
        add = TRUE
      )
    }
  }
  dev.off()

  rthal_cross_section <- create_cross_section_matrix(
    list(rthal_mat, rthal_lpme_isomap_vals, rthal_pme_vals),
    4,
    c(0.005, 0.005, 0.005),
    nrow = 2,
    lhipp = rthal
  )
  ggsave(
    paste0(dir_path, "/adni_rthal_cross_section.png"),
    plot = rthal_cross_section,
    height = 5000,
    width = 7000,
    units = "px",
    dpi = 1250
  )
}
```
```{r}
colors <- brewer.pal(3, "Set1")

roster <- read_csv(here("data/adni_standard/ROSTER.csv"))
roster <- roster %>% 
  select(
    -USERDATE,
    -USERDATE2,
    -update_stamp
  )
adni_status <- read_csv(here("data/adni_standard/ADNIMERGE.csv"))

# adni_status <- full_join(
#   adni_status,
#   roster,
#   by = c("Phase", "ID", "RID", "SITEID")
# )

sub_dirs <- result_dirs

hipp_info <- data.frame(
  patno = character(),
  date = numeric(),
  lhipp_data_vol = numeric(),
  lhipp_vol_lpme = numeric(),
  lhipp_vol_pme = numeric(),
  rhipp_data_vol = numeric(),
  rhipp_vol_lpme = numeric(),
  rhipp_vol_pme = numeric()
)

thal_info <- data.frame(
  patno = character(),
  date = numeric(),
  lthal_data_vol = numeric(),
  lthal_vol_lpme = numeric(),
  lthal_vol_pme = numeric(),
  rthal_data_vol = numeric(),
  rthal_vol_lpme = numeric(),
  rthal_vol_pme = numeric()
)

for (dir in sub_dirs) {
  file_list <- list.files(dir, full.names = FALSE)
  if ("hipp_info.csv" %in% file_list) {
    temp_hipp_info <- read_csv(paste0(dir, "/hipp_info.csv"))[, -1]
    hipp_info <- bind_rows(hipp_info, temp_hipp_info)
  }
  if ("thal_info.csv" %in% file_list) {
    temp_thal_info <- read_csv(paste0(dir, "/thal_info.csv"))[, -1]
    thal_info <- bind_rows(thal_info, temp_thal_info)
  }
}

hipp_bl <- hipp_info %>%
  group_by(patno) %>%
  summarize(
    date_bl = min(date),
    max_date = max(date)
  )
thal_bl <- thal_info %>%
  group_by(patno) %>%
  summarize(
    date_bl = min(date),
    max_date = max(date)
  )

eligible_patnos <- hipp_bl %>%
  filter(max_date - date_bl > 2) %>%
  select(patno) %>%
  unlist() %>%
  unique()

hipp_info <- full_join(hipp_info, hipp_bl, by = "patno") %>%
  mutate(time_from_bl = date - date_bl) %>%
  filter(patno %in% eligible_patnos)

thal_info <- full_join(thal_info, thal_bl, by = "patno") %>%
  mutate(time_from_bl = date - date_bl) %>%
  filter(patno %in% eligible_patnos)

hipp_summary <- hipp_info %>%
  group_by(patno) %>%
  summarize(
    lhipp_data_vol_sd = sd(lhipp_data_vol),
    lhipp_vol_lpme_sd = sd(lhipp_vol_lpme),
    lhipp_vol_pme_sd = sd(lhipp_vol_pme),
    rhipp_data_vol_sd = sd(rhipp_data_vol),
    rhipp_vol_lpme_sd = sd(rhipp_vol_lpme),
    rhipp_vol_pme_sd = sd(rhipp_vol_pme)
  )

hipp_long <- hipp_info %>%
  pivot_longer(
    lhipp_data_vol:lhipp_vol_pme,
    names_to = "lhipp_method",
    values_to = "lhipp_volume"
  ) %>%
  pivot_longer(
    rhipp_data_vol:rhipp_vol_pme,
    names_to = "rhipp_method",
    values_to = "rhipp_volume"
  ) %>%
  mutate(
    lhipp_method = gsub("lhipp_", "", lhipp_method),
    rhipp_method = gsub("rhipp_", "", rhipp_method)
  ) %>%
  filter(lhipp_method == rhipp_method) %>%
  select(-rhipp_method) %>%
  separate(lhipp_method, c("source", "method"), sep = "_") %>%
  mutate(
    source_mod = ifelse(source == "data", source, method),
    method_mod = ifelse(source == "vol", source, method)
  ) %>%
  select(-source, -method) %>%
  rename(source = source_mod, method = method_mod)

thal_long <- thal_info %>%
  pivot_longer(
    lthal_data_vol:lthal_vol_pme,
    names_to = "lthal_method",
    values_to = "lthal_volume"
  ) %>%
  pivot_longer(
    rthal_data_vol:rthal_vol_pme,
    names_to = "rthal_method",
    values_to = "rthal_volume"
  ) %>%
  mutate(
    lthal_method = gsub("lthal_", "", lthal_method),
    rthal_method = gsub("rthal_", "", rthal_method)
  ) %>%
  filter(lthal_method == rthal_method) %>%
  select(-rthal_method) %>%
  separate(lthal_method, c("source", "method"), sep = "_") %>%
  mutate(
    source_mod = ifelse(source == "data", source, method),
    method_mod = ifelse(source == "vol", source, method)
  ) %>%
  select(-source, -method) %>%
  rename(source = source_mod, method = method_mod)

thal_info_ts <- as_tsibble(
  thal_info,
  key = patno,
  index = time_from_bl,
  regular = FALSE
)

hipp_info_ts <- as_tsibble(
  hipp_info,
  key = patno,
  index = time_from_bl,
  regular = FALSE
)

hipp_info_ts %>%
  sample_n_keys(3) %>%
  ggplot(aes(x = time_from_bl, group = patno, color = patno)) +
  geom_line(aes(y = lhipp_data_vol), linetype = "solid") +
  geom_line(aes(y = lhipp_vol_lpme), linetype = "dashed") +
  geom_line(aes(y = lhipp_vol_pme), linetype = "dotted") +
  xlab("Time from Baseline Visit (Years)") +
  ylab("Estimated Left Hippocampus Volume")

lhipp_data_vol_pred <- vector()
lhipp_lpme_vol_pred <- vector()
lhipp_pme_vol_pred <- vector()
rhipp_data_vol_pred <- vector()
rhipp_lpme_vol_pred <- vector()
rhipp_pme_vol_pred <- vector()

for (patno_val in unique(hipp_info$patno)) {
  patno_data <- hipp_info %>%
    filter(patno == patno_val)
  lhipp_data_lm <- lm(lhipp_data_vol ~ date, data = patno_data)
  lhipp_data_pred <- predict(lhipp_data_lm, newdata = patno_data)
  lhipp_data_vol_pred <- c(lhipp_data_vol_pred, lhipp_data_pred)

  lhipp_lpme_lm <- lm(lhipp_vol_lpme ~ date, data = patno_data)
  lhipp_lpme_pred <- predict(lhipp_lpme_lm, newdata = patno_data)
  lhipp_lpme_vol_pred <- c(lhipp_lpme_vol_pred, lhipp_lpme_pred)

  lhipp_pme_lm <- lm(lhipp_vol_pme ~ date, data = patno_data)
  lhipp_pme_pred <- predict(lhipp_pme_lm, newdata = patno_data)
  lhipp_pme_vol_pred <- c(lhipp_pme_vol_pred, lhipp_pme_pred)

  rhipp_data_lm <- lm(rhipp_data_vol ~ date, data = patno_data)
  rhipp_data_pred <- predict(rhipp_data_lm, newdata = patno_data)
  rhipp_data_vol_pred <- c(rhipp_data_vol_pred, rhipp_data_pred)

  rhipp_lpme_lm <- lm(rhipp_vol_lpme ~ date, data = patno_data)
  rhipp_lpme_pred <- predict(rhipp_lpme_lm, newdata = patno_data)
  rhipp_lpme_vol_pred <- c(rhipp_lpme_vol_pred, rhipp_lpme_pred)

  rhipp_pme_lm <- lm(rhipp_vol_pme ~ date, data = patno_data)
  rhipp_pme_pred <- predict(rhipp_pme_lm, newdata = patno_data)
  rhipp_pme_vol_pred <- c(rhipp_pme_vol_pred, rhipp_pme_pred)
}

hipp_lm_pred <- tibble(
  lhipp_data_vol_pred,
  lhipp_lpme_vol_pred,
  lhipp_pme_vol_pred,
  rhipp_data_vol_pred,
  rhipp_lpme_vol_pred,
  rhipp_pme_vol_pred
)

hipp_info <- bind_cols(hipp_info, hipp_lm_pred)

hipp_info_sd <- hipp_info %>%
  group_by(patno) %>%
  summarize(
    lhipp_data_sd = sd(lhipp_data_vol),
    lhipp_lpme_sd = sd(lhipp_vol_lpme),
    lhipp_pme_sd = sd(lhipp_vol_pme),
    rhipp_data_sd = sd(rhipp_data_vol),
    rhipp_lpme_sd = sd(rhipp_vol_lpme),
    rhipp_pme_sd = sd(rhipp_vol_pme),
    lhipp_data_adj_sd = sd(lhipp_data_vol - lhipp_data_vol_pred),
    lhipp_lpme_adj_sd = sd(lhipp_vol_lpme - lhipp_lpme_vol_pred),
    lhipp_pme_adj_sd = sd(lhipp_vol_pme - lhipp_pme_vol_pred),
    rhipp_data_adj_sd = sd(rhipp_data_vol - rhipp_data_vol_pred),
    rhipp_lpme_adj_sd = sd(rhipp_vol_lpme - rhipp_lpme_vol_pred),
    rhipp_pme_adj_sd = sd(rhipp_vol_pme - rhipp_pme_vol_pred)
  ) %>%
  ungroup()

hipp_sd_mean <- hipp_info_sd %>%
  summarize(
    lhipp_data_sd_mean = mean(lhipp_data_sd),
    lhipp_lpme_sd_mean = mean(lhipp_lpme_sd),
    lhipp_pme_sd_mean = mean(lhipp_pme_sd),
    rhipp_data_sd_mean = mean(rhipp_data_sd),
    rhipp_lpme_sd_mean = mean(rhipp_lpme_sd),
    rhipp_pme_sd_mean = mean(rhipp_pme_sd),
    lhipp_data_adj_sd_mean = mean(lhipp_data_adj_sd),
    lhipp_lpme_adj_sd_mean = mean(lhipp_lpme_adj_sd),
    lhipp_pme_adj_sd_mean = mean(lhipp_pme_adj_sd),
    rhipp_data_adj_sd_mean = mean(rhipp_data_adj_sd),
    rhipp_lpme_adj_sd_mean = mean(rhipp_lpme_adj_sd),
    rhipp_pme_adj_sd_mean = mean(rhipp_pme_adj_sd)
)

# print(hipp_sd_mean, width = Inf)

hipp_sd_med <- hipp_info_sd %>%
  summarize(
    lhipp_data_sd_med = median(lhipp_data_sd),
    lhipp_lpme_sd_med = median(lhipp_lpme_sd),
    lhipp_pme_sd_med = median(lhipp_pme_sd),
    rhipp_data_sd_med = median(rhipp_data_sd),
    rhipp_lpme_sd_med = median(rhipp_lpme_sd),
    rhipp_pme_sd_med = median(rhipp_pme_sd),
    lhipp_data_adj_sd_med = median(lhipp_data_adj_sd),
    lhipp_lpme_adj_sd_med = median(lhipp_lpme_adj_sd),
    lhipp_pme_adj_sd_med = median(lhipp_pme_adj_sd),
    rhipp_data_adj_sd_med = median(rhipp_data_adj_sd),
    rhipp_lpme_adj_sd_med = median(rhipp_lpme_adj_sd),
    rhipp_pme_adj_sd_med = median(rhipp_pme_adj_sd)
  )

print(hipp_sd_med, width = Inf)

thal_info_ts %>%
  sample_n_keys(3) %>%
  ggplot(aes(x = time_from_bl, group = patno, color = patno)) +
  geom_line(aes(y = lthal_data_vol), linetype = "solid") +
  geom_line(aes(y = lthal_vol_lpme), linetype = "dashed") +
  geom_line(aes(y = lthal_vol_pme), linetype = "dotted") +
  xlab("Time from Baseline Visit (Years)") +
  ylab("Estimated Left Thalamus Volume")

lthal_data_vol_pred <- vector()
lthal_lpme_vol_pred <- vector()
lthal_pme_vol_pred <- vector()
rthal_data_vol_pred <- vector()
rthal_lpme_vol_pred <- vector()
rthal_pme_vol_pred <- vector()

for (patno_val in unique(thal_info$patno)) {
  patno_data <- thal_info %>%
    filter(patno == patno_val)
  lthal_data_lm <- lm(lthal_data_vol ~ date, data = patno_data)
  lthal_data_pred <- predict(lthal_data_lm, newdata = patno_data)
  lthal_data_vol_pred <- c(lthal_data_vol_pred, lthal_data_pred)

  lthal_lpme_lm <- lm(lthal_vol_lpme ~ date, data = patno_data)
  lthal_lpme_pred <- predict(lthal_lpme_lm, newdata = patno_data)
  lthal_lpme_vol_pred <- c(lthal_lpme_vol_pred, lthal_lpme_pred)

  lthal_pme_lm <- lm(lthal_vol_pme ~ date, data = patno_data)
  lthal_pme_pred <- predict(lthal_pme_lm, newdata = patno_data)
  lthal_pme_vol_pred <- c(lthal_pme_vol_pred, lthal_pme_pred)

  rthal_data_lm <- lm(rthal_data_vol ~ date, data = patno_data)
  rthal_data_pred <- predict(rthal_data_lm, newdata = patno_data)
  rthal_data_vol_pred <- c(rthal_data_vol_pred, rthal_data_pred)

  rthal_lpme_lm <- lm(rthal_vol_lpme ~ date, data = patno_data)
  rthal_lpme_pred <- predict(rthal_lpme_lm, newdata = patno_data)
  rthal_lpme_vol_pred <- c(rthal_lpme_vol_pred, rthal_lpme_pred)

  rthal_pme_lm <- lm(rthal_vol_pme ~ date, data = patno_data)
  rthal_pme_pred <- predict(rthal_pme_lm, newdata = patno_data)
  rthal_pme_vol_pred <- c(rthal_pme_vol_pred, rthal_pme_pred)
}

thal_lm_pred <- tibble(
  lthal_data_vol_pred,
  lthal_lpme_vol_pred,
  lthal_pme_vol_pred,
  rthal_data_vol_pred,
  rthal_lpme_vol_pred,
  rthal_pme_vol_pred
)

thal_info <- bind_cols(thal_info, thal_lm_pred)

thal_info_sd <- thal_info %>%
  group_by(patno) %>%
  summarize(
    lthal_data_sd = sd(lthal_data_vol),
    lthal_lpme_sd = sd(lthal_vol_lpme),
    lthal_pme_sd = sd(lthal_vol_pme),
    rthal_data_sd = sd(rthal_data_vol),
    rthal_lpme_sd = sd(rthal_vol_lpme),
    rthal_pme_sd = sd(rthal_vol_pme),
    lthal_data_adj_sd = sd(lthal_data_vol - lthal_data_vol_pred),
    lthal_lpme_adj_sd = sd(lthal_vol_lpme - lthal_lpme_vol_pred),
    lthal_pme_adj_sd = sd(lthal_vol_pme - lthal_pme_vol_pred),
    rthal_data_adj_sd = sd(rthal_data_vol - rthal_data_vol_pred),
    rthal_lpme_adj_sd = sd(rthal_vol_lpme - rthal_lpme_vol_pred),
    rthal_pme_adj_sd = sd(rthal_vol_pme - rthal_pme_vol_pred)
  ) %>%
  ungroup()

thal_sd_mean <- thal_info_sd %>%
  summarize(
    lthal_data_sd_mean = mean(lthal_data_sd),
    lthal_lpme_sd_mean = mean(lthal_lpme_sd),
    lthal_pme_sd_mean = mean(lthal_pme_sd),
    rthal_data_sd_mean = mean(rthal_data_sd),
    rthal_lpme_sd_mean = mean(rthal_lpme_sd),
    rthal_pme_sd_mean = mean(rthal_pme_sd),
    lthal_data_adj_sd_mean = mean(lthal_data_adj_sd),
    lthal_lpme_adj_sd_mean = mean(lthal_lpme_adj_sd),
    lthal_pme_adj_sd_mean = mean(lthal_pme_adj_sd),
    rthal_data_adj_sd_mean = mean(rthal_data_adj_sd),
    rthal_lpme_adj_sd_mean = mean(rthal_lpme_adj_sd),
    rthal_pme_adj_sd_mean = mean(rthal_pme_adj_sd)
  )

# print(thal_sd_mean, width = Inf)

thal_sd_med <- thal_info_sd %>%
  summarize(
    lthal_data_sd_med = median(lthal_data_sd),
    lthal_lpme_sd_med = median(lthal_lpme_sd),
    lthal_pme_sd_med = median(lthal_pme_sd),
    rthal_data_sd_med = median(rthal_data_sd),
    rthal_lpme_sd_med = median(rthal_lpme_sd),
    rthal_pme_sd_med = median(rthal_pme_sd),
    lthal_data_adj_sd_med = median(lthal_data_adj_sd),
    lthal_lpme_adj_sd_med = median(lthal_lpme_adj_sd),
    lthal_pme_adj_sd_med = median(lthal_pme_adj_sd),
    rthal_data_adj_sd_med = median(rthal_data_adj_sd),
    rthal_lpme_adj_sd_med = median(rthal_lpme_adj_sd),
    rthal_pme_adj_sd_med = median(rthal_pme_adj_sd)
  )

print(thal_sd_med, width = Inf)

set.seed(8310)
sampled_patnos <- hipp_info_ts %>% 
  sample_n_keys(3) %>% 
  .$patno %>% 
  unique()

scale_factor <- 3
hipp_info_ts %>%
  filter(patno %in% sampled_patnos) %>%
  ggplot(aes(x = time_from_bl, group = patno)) +
  geom_point(aes(y = lhipp_data_vol), shape = 16, color = colors[1]) +
  geom_line(aes(y = lhipp_data_vol), linetype = "solid", color = colors[1]) +
  # geom_smooth(
  #   aes(y = lhipp_data_vol2),
  #   color = colors[1],
  #   method = "lm",
  #   se = FALSE,
  #   linetype = "dashed",
  #   linewidth = 1
  # ) +
  geom_point(aes(y = lhipp_vol_lpme), shape = 15, color = colors[2]) +
  geom_line(aes(y = lhipp_vol_lpme), linetype = "dashed", color = colors[2]) +
  # geom_smooth(
  #   aes(y = lhipp_vol_lpme2),
  #   color = colors[2],
  #   method = "lm",
  #   se = FALSE,
  #   linetype = "dashed",
  #   linewidth = 1
  # ) +
  geom_point(aes(y = lhipp_vol_pme), shape = 17, color = colors[3]) +
  geom_line(aes(y = lhipp_vol_pme), linetype = "dotted", color = colors[3]) +
  # geom_smooth(
  #   aes(y = lhipp_vol_pme2),
  #   color = colors[3],
  #   method = "lm",
  #   se = FALSE,
  #   linetype = "dashed",
  #   linewidth = 1
  # ) +
  facet_wrap(~patno) +
  xlab("Time from Baseline Visit (Years)") +
  ylab("Estimated Left Hippocampus Volume") + 
  theme(
    axis.title.x = element_text(size = scale_factor * 5),
    axis.title.y = element_text(size = scale_factor * 5),
    axis.text.x = element_text(size = scale_factor * 5),
    axis.text.y = element_text(size = scale_factor * 5),
    strip.text.x = element_text(size = scale_factor * 5)
  )
ggsave(here("paper/figures/adni_plots/adni_lhipp_volume_comp.png"), dpi = 1500)

thal_info_ts %>%
  filter(patno %in% sampled_patnos) %>% 
  ggplot(aes(x = time_from_bl, group = patno)) +
  geom_point(aes(y = lthal_data_vol), shape = 16, color = colors[1]) +
  geom_line(aes(y = lthal_data_vol), linetype = "solid", color = colors[1]) +
  # geom_smooth(
  #   aes(y = lthal_data_vol2),
  #   color = colors[1],
  #   method = "lm",
  #   se = FALSE,
  #   linetype = "dashed",
  #   linewidth = 1
  # ) +
  geom_point(aes(y = lthal_vol_lpme), shape = 15, color = colors[2]) +
  geom_line(aes(y = lthal_vol_lpme), linetype = "dashed", color = colors[2]) +
  # geom_smooth(
  #   aes(y = lthal_vol_lpme2),
  #   color = colors[2],
  #   method = "lm",
  #   se = FALSE,
  #   linetype = "dashed",
  #   linewidth = 1
  # ) +
  geom_point(aes(y = lthal_vol_pme), shape = 17, color = colors[3]) +
  geom_line(aes(y = lthal_vol_pme), linetype = "dotted", color = colors[3]) +
  # geom_smooth(
  #   aes(y = lthal_vol_pme2),
  #   color = colors[3],
  #   method = "lm",
  #   se = FALSE,
  #   linetype = "dashed",
  #   linewidth = 1
  # ) +
  facet_wrap(~patno) +
  xlab("Time from Baseline Visit (Years)") +
  ylab("Estimated Left Thalamus Volume") + 
  guides(
    color = guide_legend(title = "Participant ID"),
    shape = guide_legend(title = "Estimate Source")
  ) +
  theme(
    axis.title.x = element_text(size = scale_factor * 5),
    axis.title.y = element_text(size = scale_factor * 5),
    axis.text.x = element_text(size = scale_factor * 5),
    axis.text.y = element_text(size = scale_factor * 5),
    strip.text.x = element_text(size = scale_factor * 5)
  )
ggsave(here("paper/figures/adni_plots/adni_lthal_volume_comp.png"), dpi = 1500)

adni_status %>% 
  filter(
    PTID %in% eligible_patnos,
    VISCODE == "bl"
  ) %>% 
  group_by(DX_bl) %>% 
  tally()
```