### 2025-02-10 code/01_lpme_simulations.qmd
align with reproducibility template

---
title: "LPME Simulation Studies"
author: "Robert Zielinski"
date: "January 31, 2025"
format: html
---

This notebook includes all code needed to replicate the results of the simulations comparing performance between the Principal Manifold Estimation (PME), Longitudinal Principal Manifold Estimation (LPME), and Principal Curve/Surface (PC/PS) algorithms on data generated from a common underlying manifold, but with varying degrees of noise added independently between time points. The LPME algorithm is designed to obtain *smooth* estimates of the manifolds underlying noisy data. Consequently, if the algorithm achieves its stated goal, it will not show the best goodness of fit to the observed data in these conditions. Simulated data is useful in this setting because we can generate the data from a known underlying manifold, allowing us to measure the algorithm's success at recovering this known manifold.

```{r, echo=FALSE}
#| label: chunk_setup
knitr::opts_chunk$set(message = FALSE, warning = FALSE)
```

```{r, echo=FALSE}
#| label: load_packages
library(tidyverse)
library(plotly)
library(pracma)
library(foreach)
library(doParallel)
library(furrr)
library(progress)
library(pme)
library(princurve)
library(Rfast)

source("prinSurf_v3.R")
source("functions/simulations/simulate_data.R")
source("functions/simulations/preprocess_data.R")
source("functions/simulations/fit_models.R")
source("functions/simulations/evaluate_models.R")
source("functions/simulations/display_results.R")
source("functions/calculate_lpme_reconstructions.R")
source("functions/calculate_pme_reconstructions.R")
source("functions/estimate_volume.R")
source("functions/interior_identification.R")
```
## Simulated Manifolds

Machine learning algorithms traditionally have tradeoffs when compared to each other, meaning that an algorithm that displays strong performance for one manifold may be outperformed by another algorithm when considering a different manifold. Because of this, it is advantageous to evaluate manifold learning algorithms with several manifolds meant to test an algorithm's performance on data with a variety of attributes. The function given in the code chunk below will be used to generate data lying along the manifolds of interest.
In the function above, we consider eight different manifolds:

1. For $r \in [-3, 3]$, $f(r) = (r, \sin(r + \frac{\pi}{2}))$.
```{r, echo=FALSE}
#| label: plot_case1
set.seed(100)
sim_data <- simulate_data(
  duration = 1,
  interval = 0.25,
  case = 1,
  obs_noise = 0.1,
  amplitude_noise = 0.25,
  period_noise = 0.05,
  time_trend = "constant",
  time_change = 0,
  N = 1000
)
sim_df <- sim_data$df

plot_ly(
  data = sim_df,
  x = ~X1,
  y = ~time,
  z = ~X2,
  type = "scatter3d",
  mode = "markers",
  marker = list(size = 3, color = "black"),
  name = "Data"
) |>
  add_markers(
    x = ~X_true1,
    y = ~time,
    z = ~X_true2,
    marker = list(size = 3, color = "red"),
    name = "True Manifold"
  )
```

2. For $r \in [-3\pi, 3\pi]$, $f(r) = (r, \sin(r))$.
```{r, echo=FALSE}
#| label: plot_case2
set.seed(200)
sim_data <- simulate_data(
  duration = 1,
  interval = 0.25,
  case = 2,
  obs_noise = 0.1,
  amplitude_noise = 0.25,
  period_noise = 0.05,
  time_trend = "constant",
  time_change = 0,
  N = 1000
)
sim_df <- sim_data$df

plot_ly(
  data = sim_df,
  x = ~X1,
  y = ~time,
  z = ~X2,
  type = "scatter3d",
  mode = "markers",
  marker = list(size = 3, color = "black"),
  name = "Data"
) |>
  add_markers(
    x = ~X_true1,
    y = ~time,
    z = ~X_true2,
    marker = list(size = 3, color = "red"),
    name = "True Manifold"
  )
```

3. For $r \in [-\frac{4}{5}\pi, \frac{\pi}{2}]$, $f(r) = (\cos(r), \sin(r))$.
```{r, echo=FALSE}
#| label: plot_case3
set.seed(300)
sim_df <- simulate_data(
  duration = 1,
  interval = 0.25,
  case = 3,
  obs_noise = 0.15,
  amplitude_noise = 0.25,
  period_noise = 0.10,
  time_trend = "constant",
  time_change = 0,
  N = 1000
)
sim_df <- sim_data$df

plot_ly(
  data = sim_df,
  x = ~X1,
  y = ~time,
  z = ~X2,
  type = "scatter3d",
  mode = "markers",
  marker = list(size = 3, color = "black"),
  name = "Data"
) |>
  add_markers(
    x = ~X_true1,
    y = ~time,
    z = ~X_true2,
    marker = list(size = 3, color = "red"),
    name = "True Manifold"
  )
```

4. For $r \in [-1, 1]$, $f(r) = (r, r^2, r^3)$
```{r, echo=FALSE}
#| label: plot_case4
set.seed(400)
sim_data <- simulate_data(
  duration = 1,
  interval = 0.25,
  case = 4,
  obs_noise = 0.15,
  amplitude_noise = 0.25,
  period_noise = 0.10,
  time_trend = "constant",
  time_change = 0,
  N = 1000
)
sim_df <- sim_data$df

plot_ly(
  data = sim_df,
  x = ~X1,
  y = ~X2,
  z = ~X3,
  frame = ~time,
  type = "scatter3d",
  mode = "markers",
  marker = list(size = 3, color = "black"),
  name = "Data"
) |>
  add_markers(
    x = ~X_true1,
    y = ~X_true2,
    z = ~X_true3,
    frame = ~time,
    marker = list(size = 3, color = "red"),
    name = "True Manifold"
  )
```

5. For $r \in [0, 3\pi]$, $f(r) = (r, \cos(r), \sin(r))$

```{r, echo=FALSE}
#| label: plot_case5
set.seed(500)
sim_data <- simulate_data(
  duration = 1,
  interval = 0.25,
  case = 5,
  obs_noise = 0.15,
  amplitude_noise = 0.25,
  period_noise = 0.10,
  time_trend = "constant",
  time_change = 0,
  N = 1000
)
sim_df <- sim_data$df

plot_ly(
  data = sim_df,
  x = ~X1,
  y = ~X2,
  z = ~X3,
  frame = ~time,
  type = "scatter3d",
  mode = "markers",
  marker = list(size = 3, color = "black"),
  name = "Data"
) |>
  add_markers(
    x = ~X_true1,
    y = ~X_true2,
    z = ~X_true3,
    frame = ~time,
    marker = list(size = 3, color = "red"),
    name = "True Manifold"
  )

```

6. For $r_1 \in [-1, 1]$, $r_2 \in [-1, 1]$, $f(r_1, r_2) = (r_1, r_2, \|\mathbf{r}\|^2)$

```{r, echo=FALSE}
#| label: plot_case6
set.seed(600)
sim_data <- simulate_data(
  duration = 1,
  interval = 0.25,
  case = 6,
  obs_noise = 0.15,
  amplitude_noise = 0.25,
  period_noise = 0.10,
  time_trend = "constant",
  time_change = 0,
  N = 1000
)
sim_df <- sim_data$df

plot_ly(
  data = sim_df,
  x = ~X1,
  y = ~X2,
  z = ~X3,
  frame = ~time,
  type = "scatter3d",
  mode = "markers",
  marker = list(size = 3, color = "black"),
  name = "Data"
) |>
  add_markers(
    x = ~X_true1,
    y = ~X_true2,
    z = ~X_true3,
    frame = ~time,
    marker = list(size = 3, color = "red"),
    name = "True Manifold"
  )

```

7. For $r_1 \in [0, 3\pi]$, $r_2 \in [-1, 1]$, $f(r_1, r_2) = (r_1\cos(r_1), r_1\sin(r_1), r_2)$

```{r, echo=FALSE}
#| label: plot_case7
set.seed(700)
sim_data <- simulate_data(
  duration = 1,
  interval = 0.25,
  case = 7,
  obs_noise = 0.15,
  amplitude_noise = 0.05,
  period_noise = 0.01,
  time_trend = "constant",
  time_change = 0,
  N = 1000
)
sim_df <- sim_data$df

plot_ly(
  data = sim_df,
  x = ~X1,
  y = ~X2,
  z = ~X3,
  frame = ~time,
  type = "scatter3d",
  mode = "markers",
  marker = list(size = 3, color = "black"),
  name = "Data"
) |>
  add_markers(
    x = ~X_true1,
    y = ~X_true2,
    z = ~X_true3,
    frame = ~time,
    marker = list(size = 3, color = "red"),
    name = "True Manifold"
  )

```

8. For $r_1 \in [0, \pi]$, $r_2 \in [0, 2\pi]$, $f(r_1, r_2) = (\sin(r_1)\cos(r_2), \sin(r_1)\sin(r_2), \cos(r_1))$

```{r, echo=FALSE}
#| label: plot_case8
set.seed(800)
sim_data <- simulate_data(
  duration = 1,
  interval = 0.25,
  case = 8,
  obs_noise = 0.15,
  amplitude_noise = 0.25,
  period_noise = 0.05,
  time_trend = "constant",
  time_change = 0,
  N = 1000
)
sim_df <- sim_data$df

plot_ly(
  data = sim_df,
  x = ~X1,
  y = ~X2,
  z = ~X3,
  frame = ~time,
  type = "scatter3d",
  mode = "markers",
  marker = list(size = 3, color = "black"),
  name = "Data"
) |>
  add_markers(
    x = ~X_true1,
    y = ~X_true2,
    z = ~X_true3,
    frame = ~time,
    marker = list(size = 3, color = "red"),
    name = "True Manifold"
  )

```

Note that case 8, in which observations drawn from a 2-dimensional manifold take the form of a sphere in 3 dimensions, is of particular interest due to our application to the surfaces of subcortical structures in the brain.

Simulation runs were conducted by varying the following arguments:

- `duration`: The simulated study's follow-up duration. Values tested are 1, 2, and 5, representing study durations of 1, 2, and 5 years, respectively.
- `interval`: The scheduled interval of study visits. Values tested are 0.25 and 0.5, representing study visits every 3 months and every 6 months, respectively.
- `case`: Specifies the manifold used to generate the data. 
- `obs_noise`: The standard deviation of the independent Gaussian observation-level noise. Values tested include {0.05, 0.1, 0.25, 0.5}.
- `amplitude_noise`: The standard deviation of Gaussian noise applied to the amplitude of the manifold at each time point. Values tested include {0.05, 0.1, 0.25, 0.5}.
- `period_noise`: The standard deviation of Gaussian noise applied to the period of the manifold at each time point. Values tested include {0.05, 0.1, 0.25, 0.5}.
- `time_trend`: Specifies systematic longitudinal changes in the manifold amplitude. Possible values include "constant", indicating that the manifold does not change over time, "linear", representing a constant rate of change in manifold amplitude, and "quadratic", indicating a quadratic rate of change in the manifold amplitude. To approximate trends observed when the brain atrophies, all changes in amplitude are downward.
- `time_change`: Scales the longitudinal amplitude changes. Values tested include {0, 0.05, 0.1, 0.25, 0.5}.

The sample size at each time point is set to 1,000. A factorial design was used to specify combinations of parameter values, with each simulation case being run a total of 5,760 times.

## Simulation Analysis Pipeline

For our purposes, analysis of the simulated data consists of the following steps:

- Preprocessing: Standardize observations to fall between -1 and 1 and augment with polar or spherical coordinates if needed.
- Manifold learning: Fit the LPME, PME and Principal Curve/Surface algorithms to the observed data.
- Model evaluation: For each model, obtain reconstructed estimates of the original observations. Compute the mean squared distance between these estimates and the D-dimensional observations generated from the true manifold.

Functions used to obtain the algorithm-specific reconstructions are provided in the code chunk below:

Functions specifying each step of the analysis process can be found at `code/functions/simulations/preprocess_data.R`, `code/functions/simulations/fit_models.R`, and `code/functions/simulations/evaluate_models.R`. An additional function used to generate plots to display the resulting model estimates can be found at `code/functions/simulations/display_results.R`.

Demonstrations of this analysis pipeline for several simulation cases can be found below.

### Simulation Demonstrations

```{r, eval=FALSE}
#| label: demo_case1
set.seed(100)
sim_data <- simulate_data(
  duration = 1,
  interval = 0.25,
  case = 1,
  obs_noise = 0.1,
  amplitude_noise = 0.25,
  period_noise = 0.05,
  time_trend = "constant",
  time_change = 0,
  N = 1000
)
sim_df <- sim_data$df

sim_processed <- preprocess_data(sim_df, case = 1, d = 1, D = 2)
sim_models <- fit_models(sim_processed, case = 1, d = 1, D = 2)
model_errors <- evaluate_models(
  sim_processed,
  sim_models,
  case = 1,
  d = 1,
  D = 2
)
model_errors
result_plot <- display_results(
  sim_processed,
  sim_models,
  case = 1,
  d = 1,
  D = 2
)
result_plot
```

```{r, eval=FALSE}
#| label: demo_case3
set.seed(300)
sim_data <- simulate_data(
  duration = 1,
  interval = 0.25,
  case = 3,
  obs_noise = 0.1,
  amplitude_noise = 0.25,
  period_noise = 0.05,
  time_trend = "constant",
  time_change = 0,
  N = 1000
)
sim_df <- sim_data$df

sim_processed <- preprocess_data(sim_df, case = 3, d = 1, D = 2)
sim_models <- fit_models(sim_processed, case = 3, d = 1, D = 2)
model_errors <- evaluate_models(
  sim_processed,
  sim_models,
  case = 3,
  d = 1,
  D = 2
)
model_errors

result_plot <- display_results(
  sim_processed,
  sim_models,
  case = 3,
  d = 1,
  D = 2
)
result_plot
```

```{r, eval=FALSE}
#| label: demo_case7
set.seed(700)
sim_data <- simulate_data(
  duration = 1,
  interval = 0.25,
  case = 7,
  obs_noise = 0.1,
  amplitude_noise = 0.01,
  period_noise = 0.01,
  time_trend = "constant",
  time_change = 0,
  N = 1000
)
sim_df <- sim_data$df

sim_processed <- preprocess_data(sim_df, case = 7, d = 2, D = 3)
sim_models <- fit_models(sim_processed, case = 7, d = 2, D = 3)
model_errors <- evaluate_models(
  sim_processed,
  sim_models,
  case = 7,
  d = 2,
  D = 3
)
model_errors

result_plot <- display_results(
  sim_processed,
  sim_models,
  case = 7,
  d = 2,
  D = 3
)
result_plot
```

```{r, eval=FALSE}
#| label: demo_case8
set.seed(800)
sim_data <- simulate_data(
  duration = 1,
  interval = 0.25,
  case = 8,
  obs_noise = 0.1,
  amplitude_noise = 0.25,
  period_noise = 0.05,
  time_trend = "constant",
  time_change = 0,
  N = 1000
)
sim_df <- sim_data$df

sim_processed <- preprocess_data(sim_df, case = 8, d = 2, D = 3)
sim_models <- fit_models(sim_processed, case = 8, d = 2, D = 3)
model_errors <- evaluate_models(
  sim_processed,
  sim_models,
  case = 8,
  d = 2,
  D = 3
)
model_errors

result_plot <- display_results(
  sim_processed,
  sim_models,
  case = 8,
  d = 2,
  D = 3
)
result_plot
```

## Running Simulations

The code chunks below specify the combinations of simulation parameters and runs the simulations in parallel. Note that due to computational demands, these simulations are not run in the process of compiling this notebook. This was handled by the `code/run_simulations.R` script. However, the code chunk below will yield identical results, and those same results are discussed later in the notebook.

```{r, eval=FALSE}
#| label: simulation_parameters

# set possible values of simulation parameters as discussed above
durations <- c(1, 2, 5)
intervals <- c(0.25, 0.5)
cases <- 1:8
obs_noises <- c(0.05, 0.1, 0.25, 0.5)
amplitude_noises <- c(0.05, 0.1, 0.25, 0.5)
period_noises <- c(0.05, 0.1, 0.25, 0.5)
time_trends <- c("constant", "linear", "quadratic")
time_changes <- c(0, 0.05, 0.1, 0.25, 0.5)

# create dataframe of all combinations of parameter values
parameter_df <- expand_grid(
  duration = durations,
  interval = intervals,
  case = cases,
  obs_noise = obs_noises,
  amplitude_noise = amplitude_noises,
  period_noise = period_noises,
  time_trend = time_trends,
  time_change = time_changes
)

# create empty dataframe to store error estimates
error_df <- tibble(
  lpme_error = vector(mode = "numeric", length = nrow(parameter_df)),
  pme_error = vector(mode = "numeric", length = nrow(parameter_df)),
  pc_error = vector(mode = "numeric", length = nrow(parameter_df)),
  data_error = vector(mode = "numeric", length = nrow(parameter_df))
)

# save the seed state that was set when each simulation was run
seed_states <- list()
```

```{r, eval=FALSE}
#| label: run_simulations
print("Starting Simulations...")

cl <- makeCluster(parallel::detectCores() - 1)
# unclear why, but registerDoSNOW worked better when running with slurm
registerDoSNOW(cl)

# the registerDoRNG function allows reproducibility
# of the simulations when run in parallel
registerDoRNG(611492)

pb  <- txtProgressBar(max = nrow(parameter_df), style = 3)
progress <- function(n) setTxtProgressBar(pb, n)
opts <- list(progress = progress)
error_df <- foreach(
  row_idx = seq_len(nrow(parameter_df)),
  .export = c(
    "simulate_data",
    "preprocess_data",
    "fit_models",
    "evaluate_models",
    "display_results"
  ),
  .packages = c("tidyverse", "pme"),
  .combine = rbind,
  .options.snow = opts
) %dopar% {
  seed_val <- .Random.seed
  seed_states[[row_idx]] <- seed_val

  duration_value <- parameter_df$duration[row_idx]
  interval_value <- parameter_df$interval[row_idx]
  case_value <- parameter_df$case[row_idx]
  obs_noise_value <- parameter_df$obs_noise[row_idx]
  amplitude_noise_value <- parameter_df$amplitude_noise[row_idx]
  period_noise_value <- parameter_df$period_noise[row_idx]
  trend_value <- parameter_df$time_trend[row_idx]
  change_value <- parameter_df$time_change[row_idx]

  # additional stability is needed for Swiss roll and spherical cases
  # to maintain structural integrity
  if (case_value == 7) {
    amplitude_noise_value <- amplitude_noise_value / 10
    period_noise_value <- period_noise_value / 10
  } else if (case_value == 8) {
    period_noise_value <- period_noise_value / 10
  }

  D_value <- case_when(
    case_value == 1 ~ 2,
    case_value == 2 ~ 2,
    case_value == 3 ~ 2,
    case_value == 4 ~ 3,
    case_value == 5 ~ 3,
    case_value == 6 ~ 3,
    case_value == 7 ~ 3,
    case_value == 8 ~ 3
  )
  d_value <- case_when(
    case_value == 1 ~ 1,
    case_value == 2 ~ 1,
    case_value == 3 ~ 1,
    case_value == 4 ~ 1,
    case_value == 5 ~ 1,
    case_value == 6 ~ 2,
    case_value == 7 ~ 2,
    case_value == 8 ~ 2
  )

  # following the same analysis pipeline as described above
  sim_data <- simulate_data(
    duration = duration_value,
    interval = interval_value,
    case = case_value,
    obs_noise = obs_noise_value,
    amplitude_noise = amplitude_noise_value,
    period_noise = period_noise_value,
    time_trend = trend_value,
    time_change = change_value,
    N = 1000
  )
  sim_df <- sim_data$df

  sim_processed <- preprocess_data(
    sim_df,
    case = case_value,
    d = d_value,
    D = D_value
  )

  sim_models <- fit_models(
    sim_processed,
    case = case_value,
    d = d_value,
    D = D_value
  )

  model_errors <- evaluate_models(
    sim_processed,
    sim_models,
    case = case_value,
    d = d_value,
    D = D_value
  )

  result_plot <- display_results(
    sim_processed,
    sim_models,
    case = case_value,
    d = d_value,
    D = D_value
  )

  simulation_output <- list(
    data = sim_data,
    processed_data = sim_processed,
    models = sim_models,
    errors = model_errors,
    plot = result_plot,
    seed = seed_val
  )

  result_dir <- paste0("../output/simulations/case", case_value)
  if (!dir.exists(result_dir)) {
    dir.create(result_dir, recursive = TRUE)
  }

  simulation_file <- paste0(
    "duration_",
    duration_value,
    "_interval_",
    str_pad(as.character(interval_value * 100), 3, pad = "0"),
    "_obs_noise_",
    str_pad(as.character(obs_noise_value * 100), 3, pad = "0"),
    "_amplitude_noise_",
    str_pad(as.character(amplitude_noise_value * 100), 3, pad = "0"),
    "_period_noise_",
    str_pad(as.character(period_noise_value * 100), 3, pad = "0"),
    "_time_trend_",
    trend_value,
    "_time_change_",
    str_pad(as.character(change_value * 100), 3, pad = "0"),
    ".RDS"
  )

  saveRDS(simulation_output, file.path(result_dir, simulation_file))

  unlist(model_errors)
}

stopCluster(cl)
```

```{r, eval=FALSE}
#| label: save_simulation_results
error_df <- error_df %>%
  as_tibble()
print(error_df)

print("Simulations Complete")

simulation_results <- bind_cols(parameter_df, error_df)
write_csv(simulation_results, "../output/simulation_results.csv")
```
## Simulation Results

```{r}
#| label: sample_result
result_case1 <- readRDS(
  file.path(
    "../output/simulations/case1",
    "duration_1_interval_025_obs_noise_005_amplitude_noise_010_period_noise_010_time_trend_linear_time_change_010.RDS"
    # "duration_1_interval_050_obs_noise_010_amplitude_noise_025_period_noise_010_time_trend_linear_time_change_010.RDS"
    # "duration_1_interval_025_obs_noise_005_amplitude_noise_005_period_noise_005_time_trend_constant_time_change_025.RDS"
  )
)

case1_observed <- result_case1$processed_data$df_observed
case1_true <- result_case1$processed_data$df_true
case1_sim <- result_case1$data$df

lpme_data <- result_case1$models$lpme$reconstructions
pme_data <- result_case1$models$pme$reconstructions
pc_data <- result_case1$models$pc$reconstructions

case1_names <- c("time", "X1", "X2")
names(case1_observed) <- case1_names
names(case1_true) <- case1_names
case1_observed$source <- "Data"
case1_observed$param <- case1_sim$r1
case1_true$source <- "True Manifold"
case1_true$param <- case1_sim$r1

lpme_df <- as_tibble(lpme_data, ..name_repair = NULL)
names(lpme_df) <- case1_names
lpme_df$source <- "LPME"
lpme_df$param <- case1_sim$r1

pme_df <- as_tibble(pme_data, ..name_repair = NULL)
names(pme_df) <- case1_names
pme_df$source <- "PME"
pme_df$param <- case1_sim$r1

pc_df <- as_tibble(pc_data, ..name_repair = NULL)
names(pc_df) <- case1_names
pc_df$source <- "PC"
pc_df$param <- case1_sim$r1

result_case1$plot
```

As an example simulation result, we consider the run using the first manifold case with a study duration of 1 year, visits every 3 months, observation noise = 0.025, amplitude noise = 0.005, period noise = 0.005, and no change in the underlying over time. The model estimates are shown in 3 dimensions in the interactive plot above, or in 2 dimensions below. Visual ispection indicates that the PME estimate, shown in green, tends to follow the observed data at each time point, while the LPME estimate appears to fall between the PME estimate and the observed manifold at most time points. When `time = 1`, the LPME estimate is much further from the true manifold, as there are no further time points to offset the error in the observations collected at that visit.

```{r, echo=FALSE}
#| label: result_plot
case1_data <- bind_rows(
  case1_observed,
  case1_true,
  lpme_df,
  pme_df,
  pc_df
) |>
  arrange(param)

color_values <- c("black", "blue", "purple", "green", "red")

set.seed(8195)
# subsample data to improve clarity of visualization
data_sampled <- case1_data |>
  group_by(time, source) |>
  sample_n(100)
```

```{r, echo=FALSE}
#| label: result_facet_plot
ggplot(case1_data, aes(x = X1, y = X2)) +
  facet_wrap(~time, nrow = 3) +
  geom_point(
    data = subset(case1_data, source == "Data"),
    color = "black",
    alpha = 0.2,
    size = 0.25
  ) +
  geom_point(
    data = subset(case1_data, source != "Data"),
    aes(color = source, shape = source),
    size = 0.2
  ) +
  scale_color_manual(values = color_values[-1])
```

```{r}
result_case1$errors
```

Strikingly, the principal curve algorithm does not appear to fit the data well at many points, visibly losing the structure of the underlying manifold at many points.

## Atrophy Estimation

The proposed use case for the LPME algorithm is to obtain smooth estimates of the manifold underlying the shape of the hippocampus over time, which can in turn be used to estimate the volume of the hippocampus, giving insight into atrophy that occurs as Alzheimer's disease progresses. To understand how well this approach performs, we consider simulation case 8, in which data from a 2-dimensional manifold is embedded in a spherical form.

Analogous to the proposed application, the LPME estimate can be used to approximate the volume of the simulated sphere. Unlike the application, the volume of the true sphere is known and provides a ground truth, allowing us to evaluate the performance of the smoothing in the presence of data generating patterns commonly found in neuroimaging data.

```{r}
dir_case8 <- "../output/simulations/case8"
files_case8 <- list.files(dir_case8)

file_name <- files_case8[1]
file_stripped <- str_replace(file_name, ".RDS", "")
file_parsed <- str_split(file_stripped, "_")

duration_val <- as.numeric(file_parsed[[1]][2])
interval_val <- as.numeric(file_parsed[[1]][4]) / 100
time_trend_val <- file_parsed[[1]][16]
time_change_val <- as.numeric(file_parsed[[1]][19]) / 100

sim <- readRDS(file.path(dir_case8, file_name))

sim_data <- sim$processed_data$df
time_points <- sim_data |>
  select(time) |>
  unlist() |>
  unique()

sim_mat <- sim$processed_data$df_observed |>
  select(time, contains("X")) |>
  as.matrix()

radii <- map(
  seq_along(sim$data$amplitude_values),
  ~ sim$data$amplitude_values[[.x]][1]
) |>
  reduce(c)

observed_volumes <- (4 / 3) * pi * radii^3

time_adjustments <- case_when(
  time_trend_val == "constant" ~ 0 * time_points,
  time_trend_val == "linear" ~ time_points,
  time_trend_val == "quadratic" ~ time_points^2
)

# scale time adjustments to be proportion of 1
if (max(time_adjustments) == 0) {
  time_adjustments_scaled <- time_adjustments * time_change_val * 0
} else {
  time_adjustments_scaled <- (
    time_adjustments / max(time_adjustments)
  ) * time_change_val
}
# subtract from initial amplitude to replicate effects of atrophy
true_radii <- 1 - time_adjustments_scaled
true_volumes <- (4 / 3) * pi * true_radii^3
```

```{r}
lpme_interior_points <- matrix(ncol = ncol(sim_mat))
lpme_volume <- vector(mode = "numeric", length = length(time_points))
pme_interior_points <- matrix(ncol = ncol(sim_mat))
pme_volumes <- vector(mode = "numeric", length = length(time_points))

df_observed <- sim$processed_data$df_observed |>
  select(time, contains("X")) |>
  as.matrix()
df_true <- sim$processed_data$df_true |>
  select(time, contains("X")) |>
  as.matrix()
df_lpme <- sim$models$lpme$reconstructions[, 1:4]
df_pme <- sim$models$pme$reconstructions[, 1:4]

for (time_idx in seq_along(time_points)) {
  temp_observed <- df_observed[df_observed[, 1] == time_points[time_idx], -1]
  temp_true <- df_true[df_true[, 1] == time_points[time_idx], -1]
  temp_lpme <- df_lpme[df_lpme[, 1] == time_points[time_idx], -1]
  temp_pme <- df_pme[df_pme[, 1] == time_points[time_idx], -1]

  scaling_factor <- 0.01
  temp_observed_scaled <- round(temp_observed / scaling_factor)
  temp_true_scaled <- round(temp_true / scaling_factor)
  temp_lpme_scaled <- round(temp_lpme / scaling_factor)
  temp_pme_scaled <- round(temp_pme / scaling_factor)
}
```

```{r, eval=FALSE}
lpme_interior_points <- matrix(ncol = ncol(sim_mat))
lpme_volumes <- vector(mode = "numeric", length = length(time_points))

pme_interior_points <- matrix(ncol = ncol(sim_mat))
pme_volumes <- vector(mode = "numeric", length = length(time_points))

for (time_idx in seq_along(time_points)) {
  temp_data <- sim_mat[sim_mat[, 1] == time_points[time_idx], -1]

  data_limits <- colMinsMaxs(temp_data)

  limit_scaler <- 0.1
  x_seq <- seq(
    data_limits[1, 1] - abs(data_limits[1, 1] * limit_scaler),
    data_limits[2, 1] + abs(data_limits[2, 1] * limit_scaler),
    by = 0.1
  )
  y_seq <- seq(
    data_limits[1, 2] - abs(data_limits[1, 2] * limit_scaler),
    data_limits[2, 2] + abs(data_limits[2, 2] * limit_scaler),
    by = 0.1
  )
  z_seq <- seq(
    data_limits[1, 3] - abs(data_limits[1, 3] * limit_scaler),
    data_limits[2, 3] + abs(data_limits[2, 3] * limit_scaler),
    by = 0.1
  )

  point_volume <- 0.1^3

  lpme_candidates  <- expand.grid(x = x_seq, y = y_seq, z = z_seq) |>
    as.matrix()
  pme_candidates <- expand.grid(x = x_seq, y = y_seq, z = z_seq) |>
    as.matrix()

  temp_pme <- sim$models$pme$pme[[time_idx]]
  temp_pme_embedding <- temp_pme$embedding_map
  temp_pme_coefs <- temp_pme$coefs[[which.min(temp_pme$MSD)]]
  temp_pme_params <- temp_pme$parameterization[[which.min(temp_pme$MSD)]]

  lpme_est <- sim$models$lpme$lpme
  temp_lpme_coefs <- lpme_est$sol_coef_functions[[
    which.min(lpme_est$msd)
  ]](time_points[time_idx])
  temp_lpme_params <- lpme_est$parameterization_list[[which.min(lpme_est$msd)]]
  temp_lpme_params <- temp_lpme_params[
    temp_lpme_params[, 1] == time_points[time_idx],
    -1
  ]
  lpme_n_knots <- nrow(temp_lpme_params)
  d <- ncol(temp_lpme_params)
  coef_mat <- matrix(
    temp_lpme_coefs,
    lpme_n_knots + d + 1,
    byrow = TRUE
  )

  temp_lpme_embedding <- function(r) {
    t(coef_mat[1:lpme_n_knots, ]) %*% pme::etaFunc(r, temp_lpme_params, 4 - d) +
      t(coef_mat[(lpme_n_knots + 1):(lpme_n_knots + d + 1), ]) %*%
        matrix(c(1, r), ncol = 1)
  }

  lpme_interior_voxels <- interior_identification(
    temp_lpme_embedding,
    coef_mat,
    temp_lpme_params,
    lpme_candidates,
    c(0, 0, 0)
  )
  pme_interior_voxels <- interior_identification(
    temp_pme_embedding,
    temp_pme_coefs,
    temp_pme_params,
    pme_candidates,
    c(0, 0, 0)
  )

  lpme_interior <- cbind(
    time_points[time_idx],
    lpme_candidates[lpme_interior_voxels, ]
  )
  lpme_interior_points <- rbind(lpme_interior_points, lpme_interior)
  lpme_volumes[time_idx] <- nrow(lpme_interior) * point_volume

  pme_interior <- cbind(
    time_points[time_idx],
    pme_candidates[pme_interior_voxels, ]
  )
  pme_interior_points <- rbind(pme_interior_points, pme_interior)
  pme_volumes[time_idx] <- nrow(pme_interior) * point_volume
}
```